\chapter{State Space Analysis of the CPN Web Socket Protocol}\label{chap:statespace}One of the advantages of Coloured Petri Nets is the ability to conduct statespace analysis, which can be used to obtain information about a model'sbehavioural properties, and can be used to locate errors and increaseconfidence in the correctness of the CPN model.\section{State Spaces}A state space is a directed graph where each node represents a reachable marking(a state) and each arc represents an occurring binding element (a transitionfiring with specific values bound to the variables of the transition). CPN Toolsby default generates the state space in breadth-first order. Once generated, the state space can be visualised directly in CPN Tools.Starting with the node for the initial state, one can pick a node and show allnodes that are reachable from it, and in this way explore the state spacevisually. This can be very tedious and unmanageable for complex state spaces,though, and instead it is usually better to use queries to automate the analysisbased on state spaces.	\subsection{Strongly Connected Component graph}	In graph theory, a strongly connected component (SCC) of a graph is a maximal	subgraph where all nodes are reachable from each other. An SCC graph has a node	for each SCC of the graph, connected by arcs determined by the arcs in the	underlying graph between nodes that belong to different SCCs. An	SCC graph is acyclic, and an SCC is said to be trivial if it contains only one	node.		By calculating the SCC graph of the state space, some of the further	analysis becomes simpler and faster, such as determining reachability and	cyclic behaviour.	\subsection{Application of State Spaces}		The biggest drawback of state space analysis is the size a state spaces may		become very large. The number of nodes and arcs often grows exponentially with		the number of starting system parameters. \com{Size of the midel		configuration considered?}				This can be remedied by picking smaller configurations,  that encapsulate		different parts of the system. This was necessary with the WebSocket Protocol		model, as the state space took too long to generate initially.				A variant of this is situations where tokens can be generated an		unlimited amount of times on a place, making the state space infinite. This		can be remedied by modifying the model to limit the number of simultaneous		tokens in the offending place.				\com{Diskutere mer hva som måtte gjøres med WebSocket, illustrasjon}		A model that incorporates random values is not fit for computing a state		space, as the number of possible enabled bindings are arbitrary for a given		state depending on the possible random values. Some values might not get used. Other		times there is an infinite number of possible values (like a floating point		number).				\com{utpeke forskjell på non-deterministic og random}		This can sometimes be alleviated by changing the behaviour to be		deterministic, for example by replacing the random function with a small color		set, such as an index or an integer with bounds. This lets the state space		generator create bindings for every possible value. 				For the WebSocket Protocol model, this was a problem for the masking key in		WebSocket frames, which is supposed to be a random 4-byte string. The		randomisation function was simply changed to always return four zeros.\section{State space report}Once the state space has been generated, CPN Tools lets the user save astate space report as a text document. The report is split into parts that eachdescribe different aspects about the state space.To explain each section of the state space report, a simple report for theWebSocket protocol has been generated, in a configuration where no messages areset to be sent. Thus, the only thing that will happen is that a connection willbe established. Later in the chapter we will consider more elaborateconfigurations of the WebSocket protocol.		\subsection{Statistics}	The first section describes general statistics about the state space.	\begin{lstlisting}  State Space     Nodes:  17     Arcs:   16     Secs:   0     Status: Full  Scc Graph     Nodes:  17     Arcs:   16     Secs:   0	\end{lstlisting}	This state space has 17 possible markings, with 16 enabled transition	occurrences connecting them. There is one more node than there are arcs, which	means this graph is a tree.		The |Secs| field shows that it took less than one second to calculate this	state space, while the |Status| field tells whether the report is generated	from a partial or full state space.		We also see that the Scc Graph has the same number of nodes and arcs, meaning	that there are no cycles in the state space (although this was already known	from the fact that it is a tree).		\subsection{Boundedness Properties}	This section describes the minimum and maximum number of tokens for	each place in the model, as well as the actual tokens these places can have.	The text has been reformatted and truncated for readability.	\begin{lstlisting}[language={}]  Best Integer Bounds                             Upper      Lower     ClientApplication          Active_Connection 1                             1          0          Conn_Result 1                             1          0          Connection_failed 1                             0          0          Messages_received 1                             1          1          Messages_to_be_sent 1                             0          0     .....  Best Upper Multi-set Bounds     ClientApplication          Active_Connection 1                         1`()          Conn_Result 1                         1`success          Connection_failed 1                         empty          Messages_received 1                         1`[]          Messages_to_be_sent 1                         empty     .....     ClientWebSocket          Connection_status 1                         1`CONN_OPEN     .....     ServerWebSocket          Connection_Status 1                         1`CONN_OPEN     .....  Best Lower Multi-set Bounds     ClientApplication          Active_Connection 1                         empty          Conn_Result 1                         empty          Connection_failed 1                         empty          Messages_received 1                         1`[]          Messages_to_be_sent 1                         empty     .....     ClientWebSocket          Connection_status 1                         empty     .....     ServerWebSocket          Connection_Status 1                         empty     .....	\end{lstlisting}		Many places show a lower and upper bound of 1. This shows a weakness	in the approach of using lists to facilitate ordered processing of tokens: We	can't see the actual number of tokens that are in the place, because	technically there is just a list there. 		Apart from that, we see that both the client and the server has an open	connection at some point, as the |Connection_status| place on the	|ClientWebSocket| and |ServerWebSocket| pages have both had a |CONN_OPEN|	token.		\subsection{Home Properties}	This section shows all home markings. A home marking is a marking that can	always be reached no matter where we are in the state space. \com{Vise marking}	\begin{lstlisting}[language={}]  Home Markings     [17]	\end{lstlisting}	We see that there is one such marking at node 17. From earlier we know that the	state space is a tree, and if this node is always reachable it must be a leaf	and all the other nodes must be in a chain. This tells us that there is only	one possible sequence of transitions to establish a connection. We can then	confidently say that the model works correctly with this configuration.		\subsection{Liveness Properties}	This section describes liveliness of the state space. Some of the transitions	have been omitted for readability.		\begin{lstlisting}[language={}]  Dead Markings     [17]  Dead Transition Instances     ClientApplication'Fail 1     ClientApplication'Receive_data 1     ClientApplication'Send_data 1     ClientWebSocket'Filter_messages 1     .....  Live Transition Instances     None	\end{lstlisting}		A dead marking is a marking from where no other markings can be reached.	\com{from where? for which? formuler} In other words, there are no transitions	for which there are enabled bindings, and the system is effectivley stopped.	For our example, we have a single dead marking, and it is the same as our home	marking, confirming that this is a leaf node in the tree.		We also get a listing of dead transition instances, which are transitions that	never have any enabled bindings and are thus never fired. This can be useful to	detect problems with a model, but in this example it is expected for many of	the transitions, since we are not sending any kind of messages in this configuration.		Last, there are live transition instances. A transition is live if we from any	reachable marking can find an occurrence sequence containing the transition.	Our example has no such transition, which follows trivially from the fact that	there is a dead marking.	The State Space report also contains fairness properties, but this does	not apply to our model since it contains no loops. We will not explain more	about this here, and instead refer you to \cite{cpn_book} chapter 7 for more	information.	\section{TODO:} Skrive om resten av analysene.Flette inn dette:\subsection{Error discovery}An error in the model was discovered this way, in the Unwrap and Receive module,where the pong reply was adding a new list instead of appending to the old onein outgoing messages. \com{Flere detaljer - hvordan viste fielen seg? Figur,før og etter}